Modal type theory based on the intuitionistic epistemic logic

### 2018-01-20 article.tex
small fixes

Here are some examples of derivation trees.

  \begin{prooftree}
  \AxiomC{$x : A \vdash x : A$}
  \RightLabel{$\textbf{K}_I$}
  \UnaryInfC{$x : A \vdash {\bf pure \: } x : \textbf{K} A$}
  \RightLabel{$\rightarrow_i$}
  \UnaryInfC{$\vdash (\lambda x. {\bf pure \: } x) : A \to \textbf{K}A$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$f : {\bf K} (A \to B) \vdash f : {\bf K} (A \to B)$}
    \AxiomC{$x : {\bf K} A \vdash x : {\bf K} A $}
    \AxiomC{$g : A \to B$}
    \AxiomC{$y : A$}
    \BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
    \TrinaryInfC{$f : {\bf K} (A \to B), x : {\bf K} A \vdash {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} B$}
    \UnaryInfC{$f : {\bf K} (A \to B) \vdash \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} A \to {\bf K} B$}
    \UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle f, x \rangle \:\: {\bf in } \:\: g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$}
  \end{prooftree}

\begin{prooftree}
  \AxiomC{$f : A \to B \vdash f : A \to B$}
  \UnaryInfC{$f : A \to B \vdash {\bf pure \:\:} f : {\bf K }(A \to B)$}
  \AxiomC{$x : {\bf K} A \vdash x : {\bf K} A $}
  \AxiomC{$g : A \to B$}
  \AxiomC{$y : A$}
  \BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
  \TrinaryInfC{$f : A \to B, x : {\bf K} A \vdash {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle {\bf pure \:\:} f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} B$}
  \UnaryInfC{$f : A \to B \vdash \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle {\bf pure \:\:} f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} A \to {\bf K} B$}
  \UnaryInfC{$\lambda f. \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle {\bf pure \:\:} f, x \rangle \:\: {\bf in } \:\: g y : (A \to B) \to {\bf K} A \to {\bf K} B$}
\end{prooftree}

@@

2) $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$;

3) Weak commutativity condition holds:

\xymatrix{
A \otimes \mathcal{K}B & \mathcal{K}A \otimes \mathcal{K}B & \mathcal{K}(A \otimes B) \\
\mathcal{K}B \otimes A & \mathcal{K}B \otimes \mathcal{K}A & \mathcal{K}(B \otimes A)
}
\end{defin}

@@

If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta} N$, then $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.

@@

If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta} N$, then $\Gamma \vdash N : A$ by subject reduction.

@@

If $\Gamma \vdash M : A$ and $M \rightarrow_{\eta} N$, then $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.

Similar to case with $\beta$-reduction.

@@

\begin{defin}
  Let us define an endofunctor $\mathcal{K} : \mathcal{C}(\lambda) \to \mathcal{C}(\lambda)$, such that:
  \begin{itemize}
    \item ${\bf K} : A \mapsto {\bf K}A$;
    \item ${\bf K} : [x,M] \in Hom_{\mathcal{C}(\lambda)}(A,B) \mapsto \text{fmap } f = [y, {\bf let \: pure \:} x = y \in M] \in Hom_{\mathcal{C}(\lambda)}({\bf K} A, {\bf K} B)$.
  \end{itemize}
\end{defin}

@@

Easy checking.

@@

$\begin{array}{lll}
&\eta_B \circ f = & \\
&[y, {\bf pure \:} y] \circ [x, M] = & \\
&[x, {\bf pure \:} y [y := M]] =& \\
&[x, {\bf pure \:} M]& \\
&\text{On the other hand}& \\
&\text{fmap } f \circ \eta_A = &\\
&[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = & \\
&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]]& \\
&[x, {\bf let \: pure \:} x = {\bf pure \: x} {\: \bf in \:} M]& \\
&[x, {\bf pure \:} M [x := x]]& \\
&[x, {\bf pure \:} M]&
\end{array}$

@@

@@

@@

>>>>

Here are some examples of derivable terms:

1) $\vdash (\lambda x. {\bf pure \: } x) : A \to \textbf{K}A$;

2) $\vdash \lambda f. \lambda x. {\bf let \: pure} \: \langle g, y \rangle = \langle f, x \rangle {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$

3) $\lambda f. \lambda x. {\bf let \:pure} \: \langle g, y \rangle = \langle {\bf pure \:} f, x \rangle {\: \bf in \: } g y : (A \to B) \to {\bf K} A \to {\bf K} B$

@@

2) $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$.
\end{defin}

@@

If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta \eta} N$, then $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.

@@

If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta \eta} N$, then $\Gamma \vdash N : A$ by subject reduction.

@@

@@

\begin{defin}
  Let us define an endofunctor $\mathcal{K} : \mathcal{C}(\lambda) \to \mathcal{C}(\lambda)$, such that:

1) ${\bf K} : A \mapsto {\bf K}A$;

2) ${\bf K} : [x,M] \in Hom_{\mathcal{C}(\lambda)}(A,B) \mapsto \text{fmap } f = [y, {\bf let \: pure \:} x = y {\: \bf in \:} M] \in Hom_{\mathcal{C}(\lambda)}({\bf K} A, {\bf K} B)$.

@@

Easy checking using reduction rules.

@@

$\begin{array}{lll}
&\eta_B \circ f = & \\
&[y, {\bf pure \:} y] \circ [x, M] = & \\
&[x, {\bf pure \:} y [y := M]] =& \\
&[x, {\bf pure \:} M]& \\
&\text{On the other hand}& \\
&\text{fmap } f \circ \eta_A = &\\
&[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = & \\
&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]] = & \\
&[x, {\bf let \: pure \:} x = {\bf pure \: x} {\: \bf in \:} M] = & \\
&[x, {\bf pure \:} M [x := x]] = & \\
&[x, {\bf pure \:} M]&
\end{array}$

@@

iii) $\ast_{A,B} \circ (\eta_{A} \times \eta_{B}) = \eta_{A \times B}$

Follows from ii).

@@

\begin{lemma} Weak commutativity.
\end{lemma}

\begin{proof}
  Follows from tensorial strength and commutativity of product.
\end{proof}

@@

\begin{proof}
  Follows from previous lemmas.
\end{proof}
